---
title: "Project 1"
author: "Nathan Preuss"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: show
    csl: biomed-central.csl
    df_print: paged
    fig_caption: yes
    fig_height: 6
    fig_width: 7
    number_sections: yes
    theme: journal
    toc: yes
    toc_float: yes
  word_document:
    toc: yes
    toc_depth: '4'
  pdf_document:
    df_print: kable
    fig_caption: yes
    fig_height: 6
    fig_width: 7
    highlight: tango
    toc: yes
    toc_depth: 4
#bibliography: project.bib
abstract: This project analyzes the software defects data table.  4 different algorithms were proposed to automate the process of finding a defect, and the efficacy of those algorithms is analyzed in this paper.  4 differenet measures for evaluating the algorithms were used, Accuracy, Detection Rate, False Alarm Rate, and Precision
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction to the data

## About the Data Set
The data is a collection of programming modules for NASA Spacecraft Instruments.  This code was evaluated line by line for defects, and returned true if the module had defective code, and false otherwise.  Because collecting this data by hand was time consuming, several methods of predicting the existence of a defect were applied.  We wish to see how accurate those methods of predicting defects are. 

## Summary Table 

Summary Table for Evaluating Defect Prediction Algorithms:

| | | Module Has Defects |
| --- | --- | --- | --- |
| | | False | True |
|Algorithm Predicts Defects | No | a | b |
| | Yes | c | d |

Latex formula for various evaluating techniques using a,b,c,d as shown above.

$$
\begin{eqnarray}
Accuracy: P(Algorithm Is Correct) = \frac{a+d}{a+b+c+d}\\
Detection Rate: P(Predict Defect | Module Has Defect) = \frac{d}{b+d}\\
FalseAlarmRate: P(Predict Defect | Module Has No Defect) = \frac{c}{a+c}\\
Precision: P(Module Has Defect | Predict Defect) = \frac{d}{c+d}\\
\end{eqnarray}
$$

# R functions
R functions based on the above equations.

```{r accuracy}
acc=function(a,b,c,d)
{
  return ((a+d)/(a+b+c+d))
}
```

```{r detecton}
detect=function(b,d)
{
  return ((b)/(b+d))
}
```

```{r alarm}
falarm=function(a,c)
{
  return ((c)/(a+c))
}
```

```{r precision}
prec=function(c,d)
{
  return ((d)/(c+d))
}
```


# Creating Tables for the 4 different algorithms

```{r }
swd=read.csv("SWDEFECTS.csv")
head(swd)
tab=with(swd, table(predict.loc.50,defect))
barplot(tab, beside=TRUE, leg=TRUE, main="PRED_LOC * DEFECT Crosstabulation")
tab2=addmargins(tab)
tab2
```
```{r }
tab=with(swd, table(predict.evg.14.5,defect))
barplot(tab, beside=TRUE, leg=TRUE, main="PRED_EVG * DEFECT Crosstabulation")
tab2=addmargins(tab)
tab2
```

```{r }
tab=with(swd, table(predict.ivg.9.2,defect))
barplot(tab, beside=TRUE, leg=TRUE, main="PRED_IVG * DEFECT Crosstabulation")
tab2=addmargins(tab)
tab2
```

```{r }
tab=with(swd, table(predict.vg.10,defect))
barplot(tab, beside=TRUE, leg=TRUE, main="PRED_VG * DEFECT Crosstabulation")
tab2=addmargins(tab)
tab2
```


# Probability Measures for Evaluating Defect Prediction Algorithms

```{r}
#creating the different vectors that make up the table
name <- c("Lines of Code", "Cyclomatic Complexity", "Essential Complexity", "Design Complexity")
accuracy <- c(acc(400, 29, 49, 20), acc(397, 35, 52, 14), acc(441, 47, 8 , 2), acc(422, 38, 27, 11))
detection <- c(detect(29, 20), detect(35, 14), detect(47, 2), detect(38, 11))
falseAlarm <- c(falarm(400, 49), falarm(397, 52), falarm(441, 8), falarm(422, 27))
precision <- c(prec(49, 20), prec(52,14), prec(8, 2), prec(27, 11))

tab3 = data.frame(name=name, accuracy=accuracy, detection=detection, falseAlarm=falseAlarm, precision=precision)
tab3[,-1] <-round(tab3[,-1],3)
tab3
```

functions mybar() input variables:
>
    1. tab ( this will be a n by m table) - like `tab3`
    2. acc ( accuracy of the decimal output ) -- like `acc=4` for example
    
```{r mybar}
mybar=function(tab, acc)
{
  # see, it was easier to do the table as a dataframe in the previous question.
  # so I guess this outputs a dataframe and not a true table.  
  barplot(tab, beside=TRUE, leg=TRUE)
  as.data.frame(tab)
  tab[,-1] <-round(tab[,-1],acc)
  tab
}
```
    
mybar() function outputs:
>
    1. A barplot of the table
    2. Commandline output in the form of a list containing the table
    
```{r}
# Example tests of the function
# creates the barplot, no decimals for the accuracy to round

tab=with(swd, table(predict.ivg.9.2,defect))
mybar(tab, 3)
```
    
