---
title: "R package"
author: "Nathan Preuss"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction



Congratulations for getting this far. To complete the evidence that you have made a working R package please answer the following by using the supplied rmd  and upload the knitted html to CANVAS.

I will assume your package is called **MATH4753PREU0001** and you have packaged and documented functions `chebyshev`, `nathancltp`, `bindif` and `fitCompare`.

Your actual package name and functions will obviously be called something else.

# First function

Please place your first function in the r chunk in the following way:

Once you are ready please remove the option `eval = FALSE`

```{r first}
MATH4753PREU0001::chebyshev(2, c(1,2,3,4,5,6,7,8,9,11,13,16,18,21))
MATH4753PREU0001::chebyshev(1, c(1,2,3,11,13,16,18,21,34,41,56))

```


# Second function

Please place your second function in the r chunk in the following way:

Once you are ready please remove the option `eval = FALSE`

```{r second}
MATH4753PREU0001::nathancltp(50, 1000, 6)
MATH4753PREU0001::nathancltp(10, 6000)
MATH4753PREU0001::nathancltp(20, 100, 2)

```


# Third function

Please place your third function in the r chunk in the following way:

Once you are ready please remove the option `eval = FALSE`

```{r third}
MATH4753PREU0001::bindif()
MATH4753PREU0001::bindif(iter=10000, n=8, p=0.2)
MATH4753PREU0001::bindif(iter = 500)
MATH4753PREU0001::bindif(n=7)
MATH4753PREU0001::bindif(p=0.5)
MATH4753PREU0001::bindif(n=8, p=0.2)

```


# Forth function

Please place your forth function in the r chunk in the following way:

Once you are ready please remove the option `eval = FALSE`

```{r forth}
x <- c(1:20)
y <- c(2:21)
z <- c(1,4,6,8,1,2,6,4,1,2,3,4,6,2,3,8,6,4,3,2)
MATH4753PREU0001::fitCompare(stats::lm(y~x),stats::lm(z~x))
MATH4753PREU0001::fitCompare(stats::lm(z~x),stats::lm(z~y))
```

Yes, there is a warning, but that's just because I gave it a model with perfect fit on purpose.
