---
title: "Lab 11"
author: "Nathan Preuss"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_folding: show
    csl: biomed-central.csl
    df_print: paged
    fig_caption: yes
    fig_height: 6
    fig_width: 7
    number_sections: yes
    theme: journal
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Task 1

Working Directory

```{r}
getwd()
```

# Task 2

## Using R as a calculator for 95, 90, 80, 50% confidence intervals for the mean

```{r}
d=c(5.0581, 4.9707, 5.0893, 4.9334, 4.9777, 5.0285, 4.8555, 4.9565, 
4.9769, 4.9722, 4.999, 4.9925, 4.9686, 5.0662, 4.9239, 4.9781, 
5.0485, 5.0014, 4.9957, 5.0195, 5.0118, 4.9928, 5.0361, 5.0185, 
4.9879)
t95=qt(0.975,24)
t90=qt(0.95,24)
t80=qt(0.90,24)
t50=qt(0.75,24)
ci=c()
ci[1]=mean(d)-t95*sd(d)/sqrt(25)
ci[2]=mean(d)+t95*sd(d)/sqrt(25)
ci[3]=mean(d)-t90*sd(d)/sqrt(25)
ci[4]=mean(d)+t90*sd(d)/sqrt(25)
ci[5]=mean(d)-t80*sd(d)/sqrt(25)
ci[6]=mean(d)+t80*sd(d)/sqrt(25)
ci[7]=mean(d)-t50*sd(d)/sqrt(25)
ci[8]=mean(d)+t50*sd(d)/sqrt(25)
ci
```
| CI for the mean | Lower Bound | Upper Bound |
| --- | ---- |----|
|95 | 4.9741 | 5.0146 | 
| 90 | 4.9776 | 5.0111 | 
| 80 | 4.9814 | 5.0072 |
| 50 | 4.9876 | 5.0011 | 

## 80% confidence interval using t.test

```{r}
obj=t.test(d,conf.level=0.80)
obj$conf.int
```

The 80% confidence interval is (4.9814, 5.0073), which is the same confidence interval as found above.

## Using R as a calculator for 95, 90, 80, 50% confidence intervals for the variance

```{r}
xsq95U=qchisq(0.975,24)
xsq95L=qchisq(0.025,24)
xsq90U=qchisq(0.95,24)
xsq90L=qchisq(0.05,24)
xsq80U=qchisq(0.90,24)
xsq80L=qchisq(0.1,24)
xsq50U=qchisq(0.75,24)
xsq50L=qchisq(0.25,24)
ci=c()
ci[1]=24*var(d)/xsq95U
ci[2]=24*var(d)/xsq95L
ci[3]=24*var(d)/xsq90U
ci[4]=24*var(d)/xsq90L
ci[5]=24*var(d)/xsq80U
ci[6]=24*var(d)/xsq80L
ci[7]=24*var(d)/xsq50U
ci[8]=24*var(d)/xsq50L
ci
```

| CI for the variance | Lower Bound | Upper Bound |
| --- | ---- |----|
|95 | .0014 | .0046 | 
| 90 | .0015 | .0041 | 
| 80 | .0017 | .0036 |
| 50 | .0020 | .0030 |

# Task 3

## 95% confidence interval for blue - snapper

```{r}
### Fish
blue=c(21.65, 17.48, 20.1, 21.57, 14.82, 19.17, 21.08, 18.23, 22.93, 
15.66, 20.89, 21.66, 18.5, 20.59, 18.63, 18.91, 19.53, 17.7, 
16.5, 19.03)

snapper=c(31.65, 27.48, 30.1, 31.57, 24.82, 29.17, 31.08, 28.23, 32.93, 
25.66, 30.89, 31.66, 28.5, 30.59, 28.63)

n1=length(blue)
n2=length(snapper)
spsq=((n1-1)*var(blue)+(n2-1)*var(snapper))/(n1+n2-2)
t=qt(0.975,n1+n2-2)
ci=c()
ci[1]=mean(blue)-mean(snapper)-t*sqrt(spsq*(1/n1+1/n2)) 
ci[2]=mean(blue)-mean(snapper)+t*sqrt(spsq*(1/n1+1/n2))
ci

#t.test(snapper,blue,conf.level=0.95,var.equal=TRUE)$conf.int

```

The 95% confidence interval for $\mu_{blue} - \mu_{snapper} = (-11.8407, -8.7575)$.

## 95% confidence interval for snapper - blue

```{r}
n1=length(snapper)
n2=length(blue)
spsq=((n1-1)*var(snapper)+(n2-1)*var(blue))/(n1+n2-2)
t=qt(0.975,n1+n2-2)
ci=c()
ci[1]=mean(snapper)-mean(blue)-t*sqrt(spsq*(1/n1+1/n2)) 
ci[2]=mean(snapper)-mean(blue)+t*sqrt(spsq*(1/n1+1/n2))
ci
```

The 95% confidence interval for $\mu_{snapper} - \mu_{blue} = (8.7575, 11.8407)$.

Given two fish, 1 snapper and 1 blue cod, we would expect to see the snapper be anywhere from 8.7575 to 11.8407 inches longer than the blue cod 95% of the time.

## 95, 85, 75, 25 ci for mean of snapper - blue

```{r}
t.test(x=snapper, y=blue, conf.level = .95, var.equal=TRUE)$conf.int
t.test(x=snapper, y=blue, conf.level = .85, var.equal=TRUE)$conf.int
t.test(x=snapper, y=blue, conf.level = .75, var.equal=TRUE)$conf.int
t.test(x=snapper, y=blue, conf.level = .25, var.equal=TRUE)$conf.int
```

| CI for the mean | Lower Bound | Upper Bound |
| --- | ---- |----|
|95 | 8.7575 | 11.8407 | 
| 85 | 9.1824 | 11.4159 | 
| 75 | 9.4119 | 11.1864 |
| 25 | 10.0557 | 10.5426 |

As the confidence level decreases, the interval gets smaller.

# Task 4

## 95% confidence interval, R-as-calculator style

```{r}
Exam1=c(40.98, 59.36, 46.69, 41.8, 61.63, 65.31, 62.96, 60.21, 56.89, 78.41, 53.44, 75.2, 60.54, 52.43, 41.41, 70.79, 73.55, 55.65, 61.43, 63.84, 58.07, 53.79, 54.45, 67.18, 44.46)
Exam2=c(50.22, 66.19, 58.75, 51.88, 66.61, 70.86, 74.25, 70.23, 69.55, 87.18, 63.62, 81.7, 70.5, 66.02, 51.35, 80.92, 85.65, 65.44, 74.37, 75.28, 67.86, 59.92, 64.42, 73.57, 57.15)

d<-mean(Exam1)-mean(Exam2)
z<- qt(.975, length(Exam1) - 1)
ci <- c()
ci[1] = d - z * sqrt(var(Exam1-Exam2)) / sqrt(length(Exam1))
ci[2] = d + z * sqrt(var(Exam1-Exam2)) / sqrt(length(Exam1))
ci
```

There is a 95% chance that a person who scored a given value on exam 1 will score 8.7102 to 10.7313 points higher on exam 2.

## Other confidence intervals, using t.test

```{r}
exdif=Exam1-Exam2
t.test(exdif, conf.level = .90)$conf.int
t.test(exdif, conf.level = .80)$conf.int
t.test(exdif, conf.level = .70)$conf.int
t.test(exdif, conf.level = .60)$conf.int
t.test(exdif, conf.level = .10)$conf.int
```

| CI for the mean | Lower Bound | Upper Bound |
| --- | ---- |----|
| 90 | -10.5584 | -8.8831 | 
| 80 | -10.3660 | -9.0755 | 
| 70 | -10.2394 | -9.2021 |
| 60 | -10.1403 | -9.0312 |
| 10 | -9.7829  | -9.6586 | 

# Task 5

## Recreate the table

```{r}
c <- c(38, 78)
d <- c(21, 7)
df <- data.frame(c, d)
df
```

## R as a calculator

Point estimate for flightless birds: $\hat p = \frac Y n = \frac {21} {38} = 0.5526$.
Point estimate for flying birds: $\hat p = \frac Y n = \frac {7} {78} = 0.0897$.

```{r}
qnorm(.975)
```

$z_{\alpha/2} = 1.9599$

Checking to make sure the sample size is sufficiently large:

$$n_1\hat p _1 = 38*0.5526 = 21.0 \ge 4$$
$$n_1\hat q _1 = 38*(1-0.5526) = 17.0 \ge 4$$
$$n_2\hat p _2 = 38*0.0897 = 7.0 \ge 4$$
$$n_2\hat q _2 = 38*(1-.0897) = 71.0 \ge 4$$

Since all of the above have values greater than 4, the normal approximation is reasonable.

$$(\hat p_1 - \hat p_2) = z_{\alpha/2} \pm \sqrt { \frac{\hat p_1 \hat q_1 } {n} + \frac{\hat p_2 \hat q_2 } {n}}$$
$$ = (.5526 - .0897) = 1.96 \pm \sqrt { \frac{.5526*(1-.5526)} {38} + \frac{.0897*(1-.0897)} {78}}$$
$$ = .4629 \pm .1703 = (.2926, .6332)$$

The confidence interval for $(\hat p_1 - \hat p_2)$ is $(.2926, .6332)$.

# Task 6

## 95% confidence interval

```{r}
set.seed(35);sam1=rnorm(25,mean=10,sd=5); 
set.seed(45);sam2=rnorm(34,mean=40,sd=8)
s1 = var(sam1)
s2 = var(sam2)
v1 = length(sam1) - 1
v2 = length(sam2) - 1
F1 = qf(.025, v1, v2)
F2 = qf(.025, v2, v1)
ci = c()
ci[2] = s1 / (F1 * s2)
ci[1] = s1 * F2 / s2
ci
```

The 95% confidence interval for $\frac{\sigma_1^2}{\sigma_2^2} = (.2257, 1.0309)$.

## Other confidence intervals

```{r}
var.test(sam1, sam2, conf.level = .8)$conf.int
var.test(sam1, sam2, conf.level = .7)$conf.int
var.test(sam1, sam2, conf.level = .6)$conf.int
var.test(sam1, sam2, conf.level = .5)$conf.int
```

| CI for the variance | Lower Bound | Upper Bound |
| --- | ---- |----|
| 80 | .2920 | .7831 | 
| 70 | .3204 | .7105 | 
| 60 | .3449 | .6581 |
| 50 | .3675 | .6165 |

# Task 7

```{r}
set.seed(23);x = rnorm(30,mean=10,sd=12)
MATH4753PREU0001::myci(x)
```

The 95% confidence interval for the mean of the sample is (8.1301, 11.8868).
