---
title: "Lab 8"
author: "Nathan Preuss"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_folding: show
    csl: biomed-central.csl
    df_print: paged
    fig_caption: yes
    fig_height: 6
    fig_width: 7
    number_sections: yes
    theme: journal
    toc: yes
    toc_float: yes
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Task 1

Working Directory

```{r}
getwd()
```

# Task 2

## Create the sample

```{r}
v <- runif(10,0,5)
```

## Theoretical Mean and Variance

$$\mu = \frac{a+b}{2} = \frac{0+5} {2} = 2.5$$
$$\sigma^2 = \frac{(b-a)^2} {12} = \frac{25}{12} = 2.0833$$

## Actual Mean and Variance

```{r}
mean(v)
var(v)
```

In this case, the actual mean was 2.8575, and the actual variance was 1.5606.  These values were not very close to the theoretical mean and variance.  Perhaps we need to increase the sample size.

## Theory Time

The expected value of $T$ is $E(T) = nE(Y_i)$.  Because the expected value and the mean are the same, this can be rewritten to be shown that the mean of $T$ depends upon $n$, the number of items in a sample, and the expected value of $y_i$ such that $\mu_T = nE(Y_i)$

Similarly, the variance of $T$ is $V(T) = nV(Y_i)$.  That is, the variance of $T$ depends upon the number of observation in a sample $n$, and the variantion for an element $Y_i$.  

$\overline Y = \frac {T} {n}$.  As a result computing its mean and variance is easy.  The expected value of $\overline Y$ is equal to the expected value of $y_i$, so the mean of $\mu_{\overline Y} = \mu_{y_i}$.  

The variance of $\overline Y$ is just the variance of $T$ divided by $n^2$, so $V(\overline Y) = \frac{V(y_i)} {n}$.

## Explaining Code

Line A creates a sample of the size $n*iter$, with $a=0$ and $b=5$ and stores this in an object called y.
Line B creates a 2D matrix with n columns, each a different sample of runif with length iter and stores this is an object called data.
Line C sums each column and stores it in an object called sm.
Line D runs the myclt function with n=10 and iter = 10000, and stores the returned object sm in W.

## Running the code

```{r}
myclt=function(n,iter){
y=runif(n*iter,0,5) # A  
data=matrix(y,nr=n,nc=iter,byrow=TRUE) #B  
sm=apply(data,2,sum) #C  
hist(sm)
sm
}  
w=myclt(n=10,iter=10000) #D
```

## Finding estimates for the sample

Looking for the sample mean and variance of w

```{r}
mean(w)
var(w)
```

The mean of w is 24.9522 and the variance is 20.7303.

```{r}
mycltrevised=function(n,iter){
y=runif(n*iter,0,5) # A  
data=matrix(y,nr=n,nc=iter,byrow=TRUE) #B  
sm=apply(data,2,sum) #C  
mn = sm/n
hist(mn)
mn
}  
w=mycltrevised(n=10,iter=10000) #D
#mean
mean(w)
#variance
var(w)
```

The mean is 2.5014 and the variance is .2091.

# Task 3

## Code explanations

The apply function uses the 2 to apply the function to the columns of the matrix.
There are iter number of items in w, so 100000.
The variance takes on that function because, as mentioned above, the variance of $\overline Y = \frac{V(y_i)} {n}$.  In this case, the $V(y_i) = \frac{(b-a)^2}{12}$.  Thus, dividing by $n$ gives us $V(\overline Y) = \frac{(b-a)^2} {n*12}$.  Then, to get from the variance to the standard deviation, we take the square root of the function, which gives us $\frac{(b-a)}{\sqrt{12n}}$, or in R: sd=(b-a)/(sqrt(12*n)). 

## Recording the Plots

```{r}
################### uniform ##########################
### CLT uniform 
## my Central Limit Function
## Notice that I have assigned default values which can be changed when the function is called
mycltu=function(n,iter,a=0,b=10){
## r-random sample from the uniform
y=runif(n*iter,a,b)
## Place these numbers into a matrix
## The columns will correspond to the iteration and the rows will equal the sample size n
data=matrix(y,nr=n,nc=iter,byrow=TRUE)
## apply the function mean to the columns (2) of the matrix
## these are placed in a vector w
w=apply(data,2,mean)
## We will make a histogram of the values in w
## How high should we make y axis?
## All the values used to make a histogram are placed in param (nothing is plotted yet)
param=hist(w,plot=FALSE)
## Since the histogram will be a density plot we will find the max density

ymax=max(param$density)
## To be on the safe side we will add 10% more to this
ymax=1.1*ymax
## Now we can make the histogram
hist(w,freq=FALSE,  ylim=c(0,ymax), main=paste("Histogram of sample mean",
"\n", "sample size= ",n,sep=""),xlab="Sample mean")
## add a density curve made from the sample distribution
lines(density(w),col="Blue",lwd=3) # add a density plot
## Add a theoretical normal curve 
curve(dnorm(x,mean=(a+b)/2,sd=(b-a)/(sqrt(12*n))),add=TRUE,col="Red",lty=2,lwd=3) # add a theoretical curve
## Add the density from which the samples were taken
curve(dunif(x,a,b),add=TRUE,lwd=4)

}
h=mycltu(n=1,iter=10000)
h=mycltu(n=2,iter=10000)
h=mycltu(n=3,iter=10000)
h=mycltu(n=5,iter=10000)
h=mycltu(n=10,iter=10000)
h=mycltu(n=30,iter=10000)
```

## Conclusion

As the sample size increases, the histogram of the sample mean approaches the histogram for the theoretical mean.

# Task 4

## Plots for p=0.3

```{r}
mycltb=function(n,iter,p=0.5,...){

## r-random sample from the Binomial
y=rbinom(n*iter,size=n,prob=p)
## Place these numbers into a matrix
## The columns will correspond to the iteration and the rows will equal the sample size n
data=matrix(y,nr=n,nc=iter,byrow=TRUE)
## apply the function mean to the columns (2) of the matrix
## these are placed in a vector w
w=apply(data,2,mean)
## We will make a histogram of the values in w
## How high should we make y axis?
## All the values used to make a histogram are placed in param (nothing is plotted yet)
param=hist(w,plot=FALSE)
## Since the histogram will be a density plot we will find the max density

ymax=max(param$density)
## To be on the safe side we will add 10% more to this
ymax=1.1*ymax

## Now we can make the histogram
## freq=FALSE means take a density
hist(w,freq=FALSE,  ylim=c(0,ymax),
main=paste("Histogram of sample mean","\n", "sample size= ",n,sep=""),
xlab="Sample mean",...)
## add a density curve made from the sample distribution
#lines(density(w),col="Blue",lwd=3) # add a density plot
## Add a theoretical normal curve 
curve(dnorm(x,mean=n*p,sd=sqrt(p*(1-p))),add=TRUE,col="Red",lty=2,lwd=3) 

}
x = mycltb(n=4,iter=10000,p=0.3)
x = mycltb(n=5,iter=10000,p=0.3)
x= mycltb(n=10,iter=10000,p=0.3)
x= mycltb(n=20,iter=10000,p=0.3)
```

## Plots for p=0.7

```{r}
x = mycltb(n=4,iter=10000,p=0.7)
x = mycltb(n=5,iter=10000,p=0.7)
x= mycltb(n=10,iter=10000,p=0.7)
x= mycltb(n=20,iter=10000,p=0.7)
```


## Plots for p=0.7

```{r}
x = mycltb(n=4,iter=10000,p=0.5)
x = mycltb(n=5,iter=10000,p=0.5)
x= mycltb(n=10,iter=10000,p=0.5)
x= mycltb(n=20,iter=10000,p=0.5)
```

## Conclusions

As the sample size increases, the histogram better approximates the theoretical histogram.  In addition, the p term appears to shift the histogram left and right on the x axis, such that an increase in p will shift the histogram right on the x axis.  For the sample size = 10, the histogram is centered at a mean of 3 when p=0.3, a mean of 5 when p = 0.5, and a mean of 7 when p = 0.7.

# Task 5

## First set of plots

```{r}
mycltp=function(n,iter,lambda=10,...){

## r-random sample from the Poisson
y=rpois(n*iter,lambda=lambda)
## Place these numbers into a matrix
## The columns will correspond to the iteration and the rows will equal the sample size n
data=matrix(y,nr=n,nc=iter,byrow=TRUE)
## apply the function mean to the columns (2) of the matrix
## these are placed in a vector w
w=apply(data,2,mean)
## We will make a histogram of the values in w
## How high should we make y axis?
## All the values used to make a histogram are placed in param (nothing is plotted yet)
param=hist(w,plot=FALSE)
## Since the histogram will be a density plot we will find the max density

ymax=max(param$density)
## To be on the safe side we will add 10% more to this
ymax=1.1*ymax

## Make a suitable layout for graphing
layout(matrix(c(1,1,2,3),nr=2,nc=2, byrow=TRUE))

## Now we can make the histogram
hist(w,freq=FALSE,  ylim=c(0,ymax), col=rainbow(max(w)),
main=paste("Histogram of sample mean","\n", "sample size= ",n," iter=",iter," lambda=",lambda,sep=""),
xlab="Sample mean",...)
## add a density curve made from the sample distribution
#lines(density(w),col="Blue",lwd=3) # add a density plot
## Add a theoretical normal curve 
curve(dnorm(x,mean=lambda,sd=sqrt(lambda/n)),add=TRUE,col="Red",lty=2,lwd=3) # add a theoretical curve

# Now make a new plot
# Since y is discrete we should use a barplot
barplot(table(y)/(n*iter),col=rainbow(max(y)), main="Barplot of sampled y", ylab ="Rel. Freq",xlab="y" )
x=0:max(y)
plot(x,dpois(x,lambda=lambda),type="h",lwd=5,col=rainbow(max(y)),
main="Probability function for Poisson", ylab="Probability",xlab="y")
}


h = mycltp(n=2,iter=10000, lambda = 4)
h = mycltp(n=3,iter=10000, lambda = 4)
h = mycltp(n=5,iter=10000, lambda = 4)
h = mycltp(n=10,iter=10000, lambda = 4)
h = mycltp(n=20,iter=10000, lambda = 4)
```


## Second set of plots

```{r}
h = mycltp(n=2,iter=10000, lambda = 10)
h = mycltp(n=3,iter=10000, lambda = 10)
h = mycltp(n=5,iter=10000, lambda = 10)
h = mycltp(n=10,iter=10000, lambda = 10)
h = mycltp(n=20,iter=10000, lambda = 10)
```

# Task 6

## Package Function and output

```{r}
MATH4753PREU0001::nathancltp(8,2000, 20)
```

