---
title: "Lab 1"
author: "Nathan Preuss"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Tasks 

## Task 1

Find the working directory

```{r}
getwd()
```

## Task 2

Read in the DDT data
```{r}
ddt <- read.csv("DDT.csv")
head(ddt)
```

## Task 3
  a. The qualitative variables are RIVER and SPECIES
  b. The quantitative variables are MILE, LENGTH, WEIGHT, DDT
  c. 3 Species
```{r}
unique(ddt[3])
```
  d.
```{r}
with(ddt, ddt[WEIGHT>800 & SPECIES=="LMBASS",])
```
  e.
```{r}
with(ddt, ddt[RIVER=="SCM" & DDT>4,])
```
  
## Clicker Questions
  a. 42.81
```{r}
summary(ddt)
```
  b.365.5461
```{r}
sapply(ddt[5], sd)
```
  c. 0 (Axis labels are reversed)
```{r}
plot(ddt$WEIGHT,ddt$LENGTH)
```
 
  d. 1.00
```{r}
v=1:20
a=v/20
tail(a)
```

## Task 4
Table of rivers
```{r}
rivertable <-table(ddt$RIVER)
rivertable
```

Color coded barplot of rivers
```{r}
rivertable <-table(ddt$RIVER)
barplot(rivertable, col=rainbow(4), xlab = "River Name", main = "Color Coded River Barplot")
```

Table of rivers crossed with species of fish
```{r}
spriv=with(ddt,table(SPECIES,RIVER))
spriv
```

Barplot of rivers crossed with species of fish
```{r}
spriv=with(ddt,table(SPECIES,RIVER))
barplot(spriv, xlab = "River")
```

## Task 5
Pie chart of Fish Species
```{r}
spec = table(ddt$SPECIES)
pie(spec, labels = names(spec))
```

Pie Chart of Rivers
```{r}
rivertable <-table(ddt$RIVER)
pie(rivertable, names(rivertable))
```

## Task 6
Boxplot of DDT

```{r}
boxplot(ddt$DDT, xlab = "DDT")
```

Boxplot of WEIGHT
```{r}
boxplot(ddt$WEIGHT, xlab = "WEIGHT")
```

Boxplot of LENGTH
```{r}
boxplot(ddt$LENGTH, xlab = "LENGTH")
```

## Task 7
Make a coplot of LENGTH Vs WEIGHT given RIVER
```{r}
coplot(LENGTH~WEIGHT|RIVER,data=ddt)
```

Make a coplot of DDT Vs WEIGHT given SPECIES
```{r}
coplot(DDT~WEIGHT|SPECIES,data=ddt)
```

## Task 8

The 3 GGPLOTS

Box Plot
```{r}
library(ggplot2)

# Number 1
x<-ggplot(ddt, aes(x=SPECIES, y=WEIGHT, fill =RIVER)) + geom_boxplot(outlier.colour = "black") + ggtitle("Nathan Preuss")
x
```

Violin Plot

```{r}
# Number 2
library(ggplot2)
y<-ggplot(ddt, aes(x=RIVER, y=LENGTH, fill =SPECIES)) + geom_violin() + ggtitle("Nathan Preuss")
y
```

Scatter Plot

```{r}
# Number 3
ddt <- read.csv("DDT.csv")
library(ggplot2)
z<-ggplot(ddt, aes(x=WEIGHT, y=LENGTH, color =SPECIES)) + geom_point() + ggtitle("Nathan Preuss")
z
```
