---
title: "Assignment 1"
author: "Nathan Preuss"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Completed 8/15 Questions

# Questions 

## Question 1

|Assignment Type | Percentage | Notes |
|---|---|---|
|Assignments | 15 | 4 of equal value|
|Laboratories | 10 | Up to 16 labs, in class exercises, Graded on canvas|
|Projects | 10 | Project 2 is worth twice that of Project 1|
|In Class Quizzes | 10 | |
|Chapter Online Canvas Quizzes | 5 ||
|Midterms | 20 ||
|Final | 30 ||

Grading Scale:

| Grade | Scores |
|---|---|
|A | 90s|
|B | 80s|
|C | 60s and 70s|
|D | 50s|
|F | <50|

Other notes: There is no alternate assessment; the grade will come from the various components listed above.  Assignments should be submitted to the appropriate dropbox on canvas.  

## Question 4
### 4(a)
The 4 types of random sampling are simple random samples, stratified random samples, cluster sampling, and systemic sampling.

### 4(b)
Simple random sampling: A sample of n experimental units from the population where each different possible sample of size n has an equal chance to be the chosen sample.

Stratified random sampling: A sample technique commonly used when the experimental units have strata, that is that characteristics of units within the strata are more similar than the characteristics between strata.  In this case, a simple random sample is taken amongst each strata, and then the random samples of the different strata are combined into one big sample for the entire data set.

Cluster sampling: In this case, experimental data are grouped into various natural clusters.  Then, various clusters are randomly sampled in order to create the sample for the data set.

Systemic sampling: In this case, every *kth* unit is chosen from a list of all experimental units to create the sample.

## Question 5
### Question 5 Textbook

The 5 wells, as asked for in the textbook.
```{r}
getwd()
mtbe <- read.csv("MTBE.csv", header=TRUE)
ind=sample(1:223,5,replace=FALSE) # random indices
mtbe[ind,]
```

### Question 5a
#### Question 5a.i

removed omitted values
```{r}
mtbeo <- na.omit(mtbe)
```

#### Question 5a.ii

calculating sd
```{r}
depth <- mtbeo[mtbeo$Aquifier == "Bedrock",]$Depth
sd(depth)
```

The standard deviation is 56.453.

## Question 8

### Question 8a
Bar graph

### Question 8b
The variable measured is types of robotics limbs

### Question 8c
The robot design currently used the most is legs ONLY.

### Question 8d

```{r}
temp <- c(15, 8, 63, 20)
freq <- temp/sum(temp)
head(freq)
```

Relative frequencies. None: 14.15%, Both: 7.54%, Legs ONLY: 59.43%, Wheels ONLY: 18.86%

### Question 8e

```{r}
pareto<-function(x,mn="Pareto barplot",...){  # x is a vector
x.tab=table(x)
xx.tab=sort(x.tab, decreasing=TRUE,index.return=FALSE)
cumsum(as.vector(xx.tab))->cs
length(x.tab)->lenx
bp<-barplot(xx.tab,ylim=c(0,max(cs)),las=2)
lb<-seq(0,cs[lenx],l=11)
axis(side=4,at=lb,labels=paste(seq(0,100,length=11),"%",sep=""),las=1,line=-1,col="Blue",col.axis="Red")
for(i in 1:(lenx-1)){
segments(bp[i],cs[i],bp[i+1],cs[i+1],col=i,lwd=2)
}
title(main=mn,...)
}
RL=c("None","Both","LegsO","WheelsO")
l=rep(RL,temp)
pareto(l)
```


## Question 10

Creating the pie chart.
```{r}
swd=read.csv("SWDEFECTS.csv", header=TRUE)
library(plotrix)
tab=table(swd$defect)
rtab=tab/sum(tab)
round(rtab,2)
pie3D(rtab,labels=list("OK","Defective"),main="pie plot of SWD")
```

Software appears to be defective about 10% of the time.

## Question 12

Putting the data into a vector and checking to see if the empirical rule applies.

```{r}
pipe <- c(1.72, 2.50, 2.16, 2.13, 1.06, 2.24, 2.31, 2.03, 1.09, 1.40, 2.57, 2.64, 1.26, 2.05, 1.19, 2.13, 1.27, 1.51, 2.41, 1.95)
hist(pipe)
```

While this distribution is not unimodal, nor is it symmetric.  Therefore, we will need to use Chebyshev's rule.
Using Chebyshev's rule with k = 4, we expect at least 15/16, or 93.75% of data points (which is about 95%) to fall within 4 standard deviations of the mean.
Calculating the standard deviation:
```{r}
sd(pipe)
mean(pipe) + 4*sd(pipe)
mean(pipe) - 4*sd(pipe)
```

Therefore, we expect about 95% of the data to fall within the range -0.215 to 3.977.

## Question 13

```{r}
ants=read.csv("test.csv", header=TRUE)
head(ants)
```

### Question 13a
```{r}
mean(ants$AntSpecies)
median(ants$AntSpecies)
```
The mean is 12.818, or the average number of ants at a site.
The median is 5, or the middle number of ants at a site.  This is much lower than the mean because there are a couple of potential outliers, where large numbers of different species of ant were found.
The mode is 4, or the most common number of species of ant found at the sites.

### Question 13b

I would recommend median as the best measure of central tendency.  The median mitigates the effects of the outliers as compared to the mean, and there are a small number of samples for the mode to be representative of the central tendency.

### Question 13c

For the Dry Steppes:
```{r}
mean(ants[1:5, 5])
median(ants[1:5, 5])
```

The mean is 40.4, the median is 40, and the mode is 40.

### Question 13d

For the Gobi Desert:
```{r}
mean(ants[6:11, 5])
median(ants[6:11, 5])
```

The mean is 28, the median is 26, and the mode is 30.

### Question 13e

Yes, the center of the plant coverage distribution appears to be different in the two different regions.  In the dry steppes, the plan coverage is higher, at about 40%, with both the mode and median reporting 40% exactly, and the mean report 40.4%.  In the desert, the plant cover is lower, with the various measures of central tendency reporting 26-30% plant cover.  Because the two different locations have a large difference in their measures of central tendency, there is a different total plant coverage distribution at the two regions.

## Question 15

Recreating the ggplot.

```{r}
ddt <- read.csv("DDT.csv")
library(ggplot2)
x<-ggplot(ddt, aes(x=RIVER, y=LENGTH, fill =SPECIES)) + geom_boxplot(outlier.colour = "black") + ggtitle("Nathan Preuss") 
#It kept giving me errors with trying to change the font to courier, so hopefully the default is courier.
x
```

