---
title: "Lab 2"
author: "Nathan Preuss"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Tasks 

## Task 1

Find the working directory

```{r}
getwd()
```

## Task 2

Read in the DDT data
```{r}
gas <- read.csv("EPAGAS.csv")
head(gas)
```

## Task 3
Make mpg a vector
```{r}
mpgVector <- as.vector(gas[,1])
class(mpgVector)
```
Calculating z
```{r}
z=(mpgVector-mean(mpgVector))/sd(mpgVector)
mean(z)
```

$$
\bar{z} = 0\\
$$

```{r}
var(z)
```

$$
s^2_z = 1
$$

Find possible outliers.  
Possible outliers have a z score between 2 and 3
```{r}
mpgVector[abs(z)>=2 & abs(z)<=3]
```

Find the actual outliers, which have a z score greater than 3

```{r}
mpgVector[abs(z)>3]
```

Latticeplot
Plotting the outliers with a different color.  Outliers are red, possible outliers are blue, and the rest are black

```{r}
library(lattice)
mycol = ifelse(abs(z)>3, "Red",
        ifelse(abs(z)>=2 &abs(z)<=3,"Blue", "Black"))  
dotplot(mpgVector,col=mycol)
```

## Task 4
Making the boxplot

```{r}
boxplot(mpgVector, notch=TRUE, col="black", main="MPG Boxplot", horizontal = TRUE)
```

With Chebyshev's rule, we would expect at least 75% of values to occur within two standard deviations of the mean

```{r}
length(mpgVector[abs(z)<2])/length(mpgVector)
```

96% of observations fall within 2 standard deviations of the mean, satisfying Chebyshev's rule

The empirical rule states that about 95% of data will fall within 2 standard deviations of the mean.  The mpg data corresponds well to this statement, as 96% of the observations fall within 2 sd instead of 95%.

The empirical rule is valid in this case, because a histograms shows a mound-shaped, symmetric distribution.

```{r}
hist(mpgVector)
```

